Containers and R---the Rockerverse and beyond

### 2020-01-14 manuscript.Rmd
add section on outsider, see #54

## Research Compendia [@benmarwick]

>>>>

## Packaging programs for processing pipelines

Docker is perfectly suited to package and execute any software and data, and is a good candidate to build complex processing pipelines independent of the used programming language.
Because of the original use case (see&nbsp;\nameref{introduction}), Docker does not have a standard mechanism to chain a number of containers together, i.e., to define how environment variables, volume mounts, or ports can be used to pass input parameters and data into a container and how to get results out.
Some packages, e.g. \pkg{containerit}, provide a Docker image that can be used similar as a CLI, but the usage is cumbersome\footnote{\href{https://o2r.info/containerit/articles/container.html}{https://o2r.info/containerit/articles/container.html}}.

**\pkg{outsider}** ([https://antonellilab.github.io/outsider/](https://antonellilab.github.io/outsider/)) tackles the problem of integrating external programmes into an R workflow \citep{bennett_outsider_2019}.
Installation and usage of external programmes can be cumbersome and even prohibited if platform dependent.
Therefore \pkg{outsider} uses the platform-independent Docker images to encapsulate processes in _outsider modules_.
Each outsider modules has a `Dockerfile` and an R package with functions for interacting with the encapsulated tool.
A user only needs to use the given functions to install a module and can then integrate a tool into their own workflow seaminglessly using only R functions.
The module takes care of transmitting arguments and sending and receiving files to and from a container.
Outsider modules are hosted, e.g., on GitHub, and \pkg{outsider} contains discovery functions for them.
<!-- illustrating outsider examples:
https://github.com/DomBennett/om..astral/blob/master/examples/astral.R
https://github.com/DomBennett/om..raxml#partitioned-dna-analysis
-->

## Research Compendia
